### 2022-09-30 reports/ACIC (2022)/poster/resources/rr.tex
Clean up and prepare for submission

% Created by tikzDevice version 0.12.3.1 on 2022-05-20 15:39:29
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

@@

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (289.08,126.47);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,126.47);
\end{scope}
\begin{scope}
\path[clip] ( 35.37, 28.21) rectangle (283.58,120.97);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37, 28.21) rectangle (283.58,120.97);
\definecolor{drawColor}{RGB}{16,36,81}

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.11,101.41) --
	(213.11,105.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.11,103.58) --
	(118.54,103.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (118.54,101.41) --
	(118.54,105.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.44, 43.43) --
	(197.44, 47.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.44, 45.61) --
	( 86.05, 45.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.05, 43.43) --
	( 86.05, 47.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.09, 72.42) --
	(212.09, 76.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.09, 74.59) --
	(104.01, 74.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.01, 72.42) --
	(104.01, 76.77);
\definecolor{fillColor}{RGB}{16,36,81}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (165.35,103.58) circle (  1.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.20, 45.61) circle (  1.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (157.25, 74.59) circle (  1.32);
\definecolor{drawColor}{RGB}{16,36,81}

\path[draw=drawColor,draw opacity=0.70,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (207.83, 28.21) -- (207.83,120.97);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37, 28.21) rectangle (283.58,120.97);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 42.85) {0.05};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 71.84) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,100.83) {0.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 45.61) --
	( 35.37, 45.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 74.59) --
	( 35.37, 74.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,103.58) --
	( 35.37,103.58);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.88, 25.46) --
	( 78.88, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.36, 25.46) --
	(143.36, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.83, 25.46) --
	(207.83, 28.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (272.30, 25.46) --
	(272.30, 28.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 78.88, 17.75) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (143.36, 17.75) {0.8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (207.83, 17.75) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (272.30, 17.75) {1.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (159.47,  7.25) {Risk ratio};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,126.47);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70, 74.59) {Hypothetical limit (mg/m$^3$)};
\end{scope}
\end{tikzpicture}

>>>>

% Created by tikzDevice version 0.12.3.1 on 2022-09-30 16:23:45
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

@@

### 2022-09-30 reports/paper/01-introduction.rmd
New revisions
Clean up and prepare for submission

Non-Hodgkin Lymphoma (NHL) is the seventh most common cancer type in the United States and has garnered intense research attention because its incidence in the United States doubled between 1973 and 1994, before plateauing at around 19 per 100,000 persons [@seer_1994;@Ekstrom-Smedby_2006]. The strongest known risk factor of NHL is immunosuppression, both congenital and acquired [@Filipovich_1992;@Chiu_2015]. Modulation of immune function contributes to the pathogenesis of NHL directly by perturbing lymphocyte function or indirectly by reducing immunosurveillance against neoplasms. Immunodeficiency may also interact with infection with Epstein-Barr Virus, resulting in uncontrolled B-cell proliferation [@Oertel_2002]. Acquired immunodeficiency through infection with human immunodeficiency virus (HIV) can lead to genetic lesions and dysregulation while increasing vulnerability to infection by other oncogenic agents [@Grulich_1999;@Alexander_2007]. However, since immunodeficiency and infection with HIV are rare, they cannot fully explain the historic rise or present burden of NHL [@Shiels_2013]. Since the rise in NHL incidence coincided with a period of rapid and extensive chemicalization in industry, agriculture and warfare, there has been stronger interest in the environmental and occupational epidemiology on NHL [@Nelson_2005;@Romero_2021]. A meta-analysis of 44 articles published between 1980 and 2014 found statistically significant associations between NHL and exposure to pesticides of several classes including carbamate, organophosporus, triazine, and organochlorine [@Schinasi_2014]. While this meta-analysis presents striking evidence of the potential explanatory role of occupational exposures, it also highlights the limited breadth of occupational exposures studied.

@@

Here, we assess the effect of hypothetical MWF exposure limits on NHL cumulative incidence in the United Auto Workers-General Motors (UAW-GM) Occupational Cohort Study. In particular, we assess the effect of limits on soluble MWF that (1) ignore exposure to other types of MWF and (2) account for other types of MWF exposure by intervening more strongly on soluble MWF when exposure to all types of MWF is high. The latter approach is more consistent with the NIOSH REL, which was recommended for exposure to any type of MWF. The large size of the study population and rich time-varying, quantitative MWF exposure data provide an opportunity to study this relatively rare cancer and evaluate realistic interventions on MWF exposure in a longitudinal cohort setting. A recent study of cancer incidence in this iconic occupational cohort found a monotonic association between NHL and exposure to soluble MWF [@Colbeth_2022]. This association was based on a traditional Cox proportional hazards regression, a method vulnerable to bias due to confounding by time-varying factors affected by prior exposure. We therefore explore this association using a novel causal estimator capable of adjusting for time-varying confounding and censoring affected by past exposure. This estimator also yields marginal effect estimates of hypothetical interventions, which are more compatible with a population health framework than conditional measures.

>>>>

Non-Hodgkin Lymphoma (NHL) is the seventh most common cancer type in the United States and has garnered intense research attention because its incidence in the United States doubled between 1973 and 1994, before plateauing at around 19 per 100,000 persons per year [@seer_1994;@Ekstrom-Smedby_2006]. The strongest known risk factor of NHL is immunosuppression, both congenital and acquired [@Filipovich_1992;@Chiu_2015]. Modulation of immune function contributes to the pathogenesis of NHL directly by perturbing lymphocyte function or indirectly by reducing immunosurveillance against neoplasms. Immunodeficiency may also interact with infection with Epstein-Barr Virus, resulting in uncontrolled B-cell proliferation [@Oertel_2002]. Acquired immunodeficiency through infection with human immunodeficiency virus (HIV) can lead to genetic lesions and dysregulation while increasing vulnerability to infection by other oncogenic agents [@Grulich_1999;@Alexander_2007]. However, since immunodeficiency and infection with HIV are rare, they cannot fully explain the historic rise or present burden of NHL [@Shiels_2013]. Since the rise in NHL incidence coincided with a period of rapid and extensive chemicalization in industry, agriculture and warfare, there has been stronger interest in the environmental and occupational epidemiology on NHL [@Nelson_2005;@Romero_2021]. A meta-analysis of 44 articles published between 1980 and 2014 found statistically significant associations between NHL and exposure to pesticides of several classes including carbamate, organophosporus, triazine, and organochlorine [@Schinasi_2014]. While this meta-analysis presents striking evidence of the potential explanatory role of occupational exposures, it also highlights the limited breadth of occupational exposures studied.

@@

Here, we assess the effect of hypothetical MWF exposure limits on NHL cumulative incidence in the United Auto Workers-General Motors (UAW-GM) Occupational Cohort Study. In particular, we assess the effect of limits on soluble MWF that (1) ignore exposure to other types of MWF and (2) account for other types of MWF exposure by intervening more strongly on soluble MWF when the sum of exposure to all types of MWF is high. The latter approach is more consistent with the NIOSH REL, which was recommended for exposure to any type of MWF. However, taking into account workers' exposure to different types of MWF may require closer monitoring and more intensive intervention. The large size of the study population and rich time-varying, quantitative MWF exposure data provide an opportunity to study this relatively rare cancer and evaluate realistic interventions on MWF exposure in a longitudinal cohort setting. A recent study of cancer incidence in this iconic occupational cohort found a monotonic association between NHL and exposure to soluble MWF [@Colbeth_2022]. This association was based on a traditional Cox proportional hazards regression, a method vulnerable to bias due to confounding by time-varying factors affected by prior exposure. We therefore explore this association using a novel estimator capable of adjusting for time-varying confounding and censoring affected by past exposure. This estimator also yields marginal effect estimates of hypothetical interventions, which are more compatible with a population health framework than conditional measures.

### 2022-09-30 reports/paper/02-methods.rmd
Clean up and prepare for submission

We estimated NHL risk from 1985 to 2005 under hypothetical limits on average annual exposure to soluble MWF by applying the novel hazard-extended iterative conditional expectation (ICE) parametric g-formula estimator [@Wen_2020]. We leveraged time-varying quantitative MWF exposure data in tandem with employment records to adjust for the HWSE. First, we estimated the expected number of NHL cases per 1000 workers that we would observe if there were no competing risks. We contrasted  this counterfactual risk to that when in addition, the hypothetical exposure limits of 0.5, 0.25, and 0.05 mg/m^3^ were enforced for soluble MWF over workers' entire working lifetimes. These hypothetical limits are static interventions because they do not consider other covariates or exposures. However, since workers were often exposed to multiple types of MWF, both daily and cumulatively, we also evaluated the effect of limiting total MWF exposure to those hypothetical limits by intervening only on exposure to soluble MWF. These interventions, which take into consideration exposure to MWF of other types, are dynamic interventions on exposure to soluble MWF. All of the hypothetical interventions we evaluate are stochastic interventions rather than deterministic interventions because post-intervention exposures take on a range of values.

@@

The UAW-GM cohort includes all hourly workers at three automobile manufacturing plants in Michigan who had worked at least 3 years by 1985. Past papers provide detailed descriptions of the cohort [@Eisen_1992;@Eisen_2001]. The present study population (N = 34,738) was restricted to the autoworkers who were at work in 1941 or not yet hired, missing no more than half of their employment history, and still alive at the start of follow-up. Autoworkers in the study population were followed for NHL incidence from January 1, 1985 until NHL diagnosis, death, 10 years after the end of employment records, or upon reaching the oldest observed age at death (108 years), whichever came earlier.

@@

Company industrial hygienists collected several hundred personal and area samples for total particulate matter (mg/m^3^) composed of MWF over many decades. Research industrial hygienists collected additional air sampling data when the cohort study was launched in the mid 1980s. These data were combined with the historical data to derive quantitative 8-hour time-weighted average exposure estimates to soluble, straight, and synthetic MWFs for each combination of job, department, and plant over time. Workers' time-weighted average annual exposure to each MWF type was determined by combining this job-exposure matrix with employment records, which recorded time-varying job type, department, and plant. For employment records that were at least half complete, gaps in the record were interpolated by carrying forward the last known job type. The exposure assessment is described in detail elsewhere [@Hallock_1994;@Woskie_1994;@Woskie_2003]. Previous analyses of NHL applied exposure lags of 1 to 20 years; we lagged cumulative MWF exposures by 10 years to account for disease latency [@Smith_2007;@Karipidis_2007;@Zhang_2019].

@@

We applied the hazard-extended ICE parametric g-formula with pooling over treatment history [@Wen_2020] to estimate 20-year counterfactual risk of non-Hodgkin lymphoma. We considered six stochastic interventions on average annual exposure to soluble MWFs after the elimination of censoring/competing risks. We considered hypothetical exposure limits starting with the NIOSH Recommended Exposure Limit (REL) for total particulate mass (PM) from MWFs (0.5 mg/m^3^). Then, we estimated the risk at half the REL (0.25 mg/m^3^), and a tenth of the REL (0.05 mg/m^3^) [@niosh_1998]. The first three interventions were static interventions which enforced these hypothetical limits on average annual exposure to soluble MWF. The remaining three interventions were dynamic interventions that reduced average annual exposure to soluble MWF with the goal of enforcing the hypothetical limits for total exposure to MWF, regardless of type. If those limits on total average annual exposure were not possible by intervention on exposure to soluble MWF alone, average annual exposure to soluble MWF was reduced to 0. Interventions were applied at hire, before the start of follow-up, through the end of follow-up. Person-years were not intervened upon if the observed average annual exposure was below the hypothetical limit. Figure 1 illustrates the static and dynamic interventions for three person-year examples. Panel A shows the observed levels of exposure to soluble MWF and to MWF of other types. Panel B shows the post-intervention levels of exposure under a static intervention capping exposure to soluble MWF at the hypothetical limit (dotted line). Panel C shows the post-intervention levels under a dynamic intervention reducing exposure to soluble MWF with the goal of capping total MWF exposure at the hypothetical limit. Note that in person-year 3, the observed level of exposure to total MWF (and thus, to soluble MWF) is below the hypothetical limit, so the post-intervention exposure is left at the observed level.

>>>>

We estimated NHL risk from 1985 to 2005 under hypothetical limits on average annual exposure to soluble MWF by applying the novel hazard-extended iterative conditional expectation (ICE) parametric g-formula estimator [@Wen_2020]. We leveraged time-varying quantitative MWF exposure data in tandem with employment records to adjust for the HWSE. First, we estimated the expected number of NHL cases per 1000 workers that we would observe if there were no competing risks. We contrasted  this counterfactual risk to that when in addition, the hypothetical exposure limits of 0.5, 0.25, and 0.05 mg/m^3^ were enforced for soluble MWF over workers' entire working lifetimes. These hypothetical limits are static interventions because they do not consider other covariates or exposures. However, since workers were often exposed to multiple types of MWF, both daily and cumulatively, we also evaluated the effect of limiting total MWF exposure to those hypothetical limits by intervening only on exposure to soluble MWF. These interventions, which take into consideration exposure to MWF of other types when deciding how much soluble MWF exposure to assign, are dynamic interventions on exposure to soluble MWF. All of the hypothetical interventions we evaluate are stochastic interventions rather than deterministic interventions because post-intervention exposures take on a range of values.

@@

The UAW-GM cohort includes all hourly workers at three automobile manufacturing plants in Michigan who had worked at least 3 years by 1985. Past papers provide detailed descriptions of the cohort [@Eisen_1992;@Eisen_2001]. The present study population (N = 34,738) was restricted to the autoworkers who were at work in 1941 or not yet hired, missing no more than half of their employment history, and still alive at the start of follow-up. Autoworkers in the study population were followed for NHL incidence from January 1, 1985 until NHL diagnosis, death, January 1, 2005 (10 years after the end of employment record availability), or age 108 years (the oldest observed age at death), whichever came earlier.

@@

Company industrial hygienists collected several hundred personal and area samples for total particulate matter (mg/m^3^) composed of MWF over many decades. Research industrial hygienists collected additional air sampling data when the cohort study was launched in the mid 1980s. These data were combined with the historical data to derive quantitative 8-hour time-weighted average exposure estimates to soluble, straight, and synthetic MWFs for each combination of job, department, and plant over time. Workers' time-weighted average annual exposure to each MWF type was determined by combining this job-exposure matrix with employment records, which recorded time-varying job type, department, and plant through 1994. For employment records that were at least half complete, gaps in the record were interpolated by carrying forward the last known job type. The exposure assessment is described in detail elsewhere [@Hallock_1994;@Woskie_1994;@Woskie_2003]. Previous analyses of NHL applied exposure lags of 1 to 20 years; we lagged cumulative MWF exposures by 10 years to account for disease latency and therefore ended follow-up on January 1, 2005 [@Smith_2007;@Karipidis_2007;@Zhang_2019].

@@

We applied the hazard-extended ICE parametric g-formula with pooling over treatment history [@Wen_2020] to estimate 20-year counterfactual risk of non-Hodgkin lymphoma. We considered six stochastic interventions on average annual exposure to soluble MWFs after the elimination of censoring/competing risks. We considered hypothetical exposure limits starting with the NIOSH Recommended Exposure Limit (REL) for total particulate mass (PM) from MWFs (0.5 mg/m^3^). Then, we estimated the risk at half the REL (0.25 mg/m^3^), and a tenth of the REL (0.05 mg/m^3^) [@niosh_1998]. The first three interventions were static interventions which enforced these hypothetical limits on average annual exposure to soluble MWF. The remaining three interventions were dynamic interventions that reduced average annual exposure to soluble MWF with the goal of enforcing the hypothetical limits for total exposure to MWF, regardless of type. 
Under these interventions, exposures in some person-years did not comply with the hypothetical limit on total exposure to MWF. That is, if the sum of exposures to straight and synthetic MWF exceeded the hypothetical limit, average annual exposure to soluble MWF was simply reduced to 0; we did not evaluate interventions on exposures to the other two classes of MWF.
Interventions were applied at hire, before the start of follow-up, through the end of follow-up. Person-years were not intervened upon if the observed average annual exposure was below the hypothetical limit. Figure 1 illustrates the static and dynamic interventions for three person-year examples. Panel A shows the observed levels of exposure to soluble MWF and to MWF of other types. Panel B shows the post-intervention levels of exposure under a static intervention capping exposure to soluble MWF at the hypothetical limit (dotted line). Panel C shows the post-intervention levels under a dynamic intervention reducing exposure to soluble MWF with the goal of capping total MWF exposure at the hypothetical limit. Note that in person-year 3, the observed level of exposure to total MWF (and thus, to soluble MWF) is below the hypothetical limit, so the post-intervention exposure is left at the observed level.

### 2022-09-30 reports/paper/03-results.rmd
Clean up and prepare for submission

Table 1 presents summary statistics of exposure and covariates for the full study population and for those diagnosed with NHL between 1985 and 2004. The cohort is predominantly white (66%) and male (87%). The median year of hire among those diagnosed with NHL was 1959 whereas the median year of hire in the full study population was almost a decade later. Age at hire was approximately the same among those with NHL and the full study population. Median lagged cumulative exposure to all three MWF types was higher among NHL cases. Soluble MWFs were the most widely used MWF type, with approximately 90% of workers ever exposed. Median cumulative exposure among the exposed was 6.5 times higher for soluble than for straight MWFs. Figure 2 shows median average annual exposure to the three MWF types among exposed workers over calendar time. Exposure to MWF generally followed a downward trend over time.
<!-- Figure 2 shows the post-intervention distribution of the logarithm of cumulative exposure to soluble MWF at end of follow-up if hypothetical limits on average annual exposure were enforced for all workers starting at hire. The interventions shifted the distribution of cumulative exposure rather than deterministically assigning exposure to a particular value. -->

@@

The observed risk of NHL over the 20-year follow-up period was 6.65 per 1000. Table 2 presents the hazard-extended ICE parametric g-formula estimates of the risk and risk ratios contrasting hypothetical limits on exposure to soluble MWF to no limit, after elimination of competing risks. The estimated risk under no limit on MWF exposure was 9.56 (8.15, 10.89) per 1000. Stronger limits on average annual exposure to soluble MWFs resulted in monotonically stronger reductions in the risk of NHL. Capping average annual exposure to soluble MWFs at 0.5 mg/m^3^, 0.25 mg/m^3^, and 0.05 mg/m^3^ resulted in a risk of 8.30 (6.52, 10.19), 8.06 (6.15, 10.15), and 7.52 (5.73, 9.51) per 1000, respectively. The risk ratios contrasting these hypothetical limits to no limit were 0.87 (0.72, 1.02), 0.84 (0.68, 1.01), and 0.79 (0.62, 0.97). Dynamic reductions in soluble exposure with the aim of capping total MWF exposure also yielded monotonically protective risk ratios: 0.84 (0.69, 0.99), 0.80 (0.64, 0.98), and 0.72 (0.48, 1.01). The dynamic interventions seeking to cap total MWF exposure at 0.5 mg/m^3^ and 0.25 mg/m^3^ attained risk reductions of similar magnitude as the static interventions enforcing (stronger) exposure limits at 0.25 and 0.05 mg/m^3^, respectively, but do so while intervening on a smaller proportion of person-years.

<!-- Results for interventions on straight and synthetic MWF are presented in Tables S1 and S2 in the Appendix. The relative risks contrasting interventions on straight and synthetic MWF exposure to the natural course were much smaller in magnitude than those for soluble; none were statistically significant. Since exposure to soluble MWF was greatest in prevalence and magnitude, limits on soluble MWF would result in interventions on a greater percentage of person-years than limits on straight or synthetic MWF. The intervention capping average annual exposure to soluble MWF at the REL affected 23.8% of the person-years in the follow-up period. Limiting straight and synthetic MWF exposure to the REL affected 3.2% and 1.1% of the person-years under follow-up, respectively. -->

>>>>

Table 1 presents summary statistics of exposure and covariates for the full study population and for those diagnosed with NHL between 1985 and 2004. The cohort is predominantly white (66%) and male (87%). The median year of hire among those diagnosed with NHL was 1959 whereas the median year of hire in the full study population was almost a decade later. Age at hire was approximately the same among those with NHL and the full study population. Median lagged cumulative exposure to all three MWF types was higher among NHL cases. Soluble MWFs were the most widely used MWF type, with approximately 90% of workers ever exposed. Median cumulative exposure among the exposed was 6.5 times higher for soluble than for straight MWFs. Figure 2 shows median average annual exposure to the three MWF types among exposed workers over calendar time. Exposure to MWF generally followed a downward trend over time.

@@

The observed risk of NHL over the 20-year follow-up period was 6.65 per 1000. Table 2 presents the hazard-extended ICE parametric g-formula estimates of the risk and risk ratios contrasting hypothetical limits on exposure to soluble MWF to no limit, after elimination of competing risks. Under an intervention eliminating competing risks, the estimated risk under no limit on MWF exposure was 9.56 (8.15, 10.89) per 1000. Stronger limits on average annual exposure to soluble MWFs resulted in monotonically stronger reductions in the risk of NHL. Capping average annual exposure to soluble MWFs at 0.5 mg/m^3^, 0.25 mg/m^3^, and 0.05 mg/m^3^ resulted in a risk of 8.30 (6.44, 10.12), 8.06 (6.00, 10.01), and 7.52 (5.57, 9.35) per 1000, respectively. The risk ratios contrasting these hypothetical limits to no limit were 0.87 (0.73, 1.02), 0.84 (0.68, 1.01), and 0.79 (0.61, 0.97). Dynamic reductions in soluble exposure with the aim of capping total MWF exposure also yielded monotonically protective risk ratios: 0.84 (0.69, 0.99), 0.80 (0.64, 0.97), and 0.72 (0.44, 0.97). The dynamic interventions seeking to cap total MWF exposure at 0.5 mg/m^3^ and 0.25 mg/m^3^ attained risk reductions of similar magnitude as the static interventions enforcing (stronger) exposure limits at 0.25 and 0.05 mg/m^3^, respectively, but do so while intervening on a smaller proportion of person-years.

### 2022-09-30 reports/paper/paper.Rmd
Clean up and prepare for submission

**Running head suggestion:** Limits on metalworking fluid for preventing NHL

**Conflicts of interests:** Authors declare no conflicts of interest

@@

**Background:** Non-Hodgkin lymphoma (NHL) incidence has increased significantly since 1960 and was recently linked with workplace exposure in the United Auto Workers-General Motors (UAW-GM) Cohort Study. Here, we investigate hypothetical interventions on workplace exposure to soluble metalworking fluids (MWF) in relation to NHL risk. Interventions on entire classes of occupational exposures rather than particular agents may result in stronger lasting protections for worker health while preventing regrettable replacements.

**Methods:** We estimated counterfactual risk of NHL from 1985 to 2004 in the UAW-GM cohort of hourly autoworkers exposed to straight, soluble, and synthetic MWF at 3 Michigan plants under hypothetical interventions. We contrasted counterfactual risk under stochastic interventions on average annual exposure to soluble MWF to that under no intervention on exposure. These stochastic interventions included (1) static interventions on soluble MWF that ignore exposure to other types and (2) dynamic interventions that seek to cap total exposure to any type of MWF by intervening on exposure to soluble MWF only. Static interventions may be more feasible as real-world policy, but dynamic interventions may result in stronger protections.

**Results:** During follow-up, 231 NHL cases occurred. Stronger hypothetical limits on average annual exposure to soluble MWF resulted in monotonically stronger reductions in NHL risk. Capping exposure to soluble MWF at 0.05 mg/m^3^ resulted in a risk ratio of 0.79 (95% CI: 0.62, 0.97).  Capping total MWF exposure at 0.05 mg/m3 by intervening only on exposure to soluble MWF resulted in a risk ratio of 0.72 (95% CI: 0.48, 1.01).

**Conclusion:** We considered realistic interventions more compatible with a population health framework than deterministic interventions and showed that stronger limits on average annual exposure to soluble MWF may confer stronger protection against NHL risk.

>>>>

**Running head suggestion:** Limits on metalworking fluid for preventing NHL.

**Conflicts of interests:** Authors declare no conflicts of interest.

@@

**Background:** Non-Hodgkin lymphoma (NHL) incidence has increased significantly since 1960 and was recently linked with workplace exposure in the United Auto Workers-General Motors (UAW-GM) Cohort Study. Here, we investigate hypothetical interventions on workplace exposure to oil- and water-based soluble metalworking fluids (MWF) in relation to NHL risk. Interventions on entire classes of occupational exposures rather than particular agents may result in stronger lasting protections for worker health.

**Methods:** We estimated counterfactual risk of NHL between 1985 and 2004 in the UAW-GM cohort of autoworkers exposed to straight, soluble, and synthetic MWF at three Michigan plants under hypothetical interventions. We estimated the effect of static and dynamic stochastic interventions on average annual exposure to soluble MWF. Unlike the static interventions, dynamic interventions take into account exposures to the other two classes of metalworking fluids, reflecting the fact that currently recommended limits apply to the total of all three exposures.

**Results:** During follow-up, 231 NHL cases occurred. Stronger hypothetical limits on average annual exposure to soluble MWF resulted in monotonically stronger reductions in NHL risk. Capping exposure to soluble MWF at 0.05 mg/m^3^ resulted in a risk ratio of 0.79 (95% CI: 0.61, 0.96). Capping total MWF exposure at 0.05 mg/m^3^ by intervening only on exposure to soluble MWF resulted in a risk ratio of 0.72 (95% CI: 0.44, 0.97).

**Conclusion:** We considered interventions on soluble MWF that address total MWF exposure levels and showed that stronger limits on average annual exposure to soluble MWF may confer stronger protection against NHL risk.
